Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-10-04 main.tex
Added balance assist in second line of abstract. Thanks Frans.

\abstract{
  Uncontrolled bicycles are generally unstable at low speeds. We add an
  automatically controlled steering motor to a consumer electric bicycle that
  stabilizes the riderless bicycle down to about 3~\si{\kph}. We hypothesize
  that a stabilized bicycle will reduce the probability of falling. To test the
  motor's possible assistance during falls, we apply varying magnitude external
  handlebar perturbations to twenty-six participants who rode on a treadmill
  with the balance assist system both activated and deactivated. The
  probability of recovering from a handlebar perturbation significantly
  increases when the balance assist is activated at a travel speed of
  6~\si{\kph}. This positive effect is most prominent at and around the
  individual riders' perturbation resistance threshold. We conclude that use of
  a balance assist system in real world bicycling would reduce the number of
  falls that occur near riders' control authority limits.
}

@@

\section*{Acknowledgements}
%
This study follows and draws from experimental and analysis methods originally
developed in unpublished research by Marco Reijne. The authors acknowledge
Felix Dauer, David Gabriel, Sierd Heida, Oliver Maier, Maarten Pelgrim, Marco
Reijne, and Arend Schwab for contributions to the development of the balance
assist bicycle. We acknowledge Shannon van de Velde and Marco Reijne's
development of the bicycle perturbation system. Allen Downey provided useful
insight in the statistical condsiderations. We thank all of the assistants that
helped run the experiments and the participants for their time an enthusiasm.
We thank Peter Stahlecker, Benjamin Gonzalez, and Riender Happee for valuable
feedback in early revisions.

>>>>

\abstract{
  Uncontrolled bicycles are generally unstable at low speeds. We add an
  automatically controlled steering motor to a consumer electric bicycle that
  stabilizes the riderless bicycle down to about 3~\si{\kph} to assist the
  rider in balancing the vehicle. We hypothesize that a stabilized bicycle will
  reduce the probability of falling. To test the motor's possible assistance
  during falls, we apply varying magnitude external handlebar perturbations to
  twenty-six participants who rode on a treadmill with the balance assist
  system both activated and deactivated. The probability of recovering from a
  handlebar perturbation significantly increases when the balance assist is
  activated at a travel speed of 6~\si{\kph}. This positive effect is most
  prominent at and around the individual riders' perturbation resistance
  threshold. We conclude that use of a balance assist system in real world
  bicycling would reduce the number of falls that occur near riders' control
  authority limits.
}

@@

\section*{Acknowledgements}
%
This study follows and draws from experimental and analysis methods originally
developed in unpublished research by Marco Reijne. The authors acknowledge
Felix Dauer, David Gabriel, Sierd Heida, Oliver Maier, Maarten Pelgrim, Marco
Reijne, and Arend Schwab for contributions to the development of the balance
assist bicycle. We acknowledge Shannon van de Velde and Marco Reijne's
development of the bicycle perturbation system. Allen Downey provided useful
insight in the statistical condsiderations. We thank all of the assistants that
helped run the experiments and the participants for their time an enthusiasm.
We thank Peter Stahlecker, Benjamin Gonzalez, and Riender Happee for valuable
feedback in early revisions. Lastly, we thank Frans van der Helm for the
consistent encouragement to execute this experiment.
